Paper

### 2025-03-14 paper/paper.qmd
better names - uniform sampling script

In practice, the approach to managing outbreaks is often reactive.
Once an outbreak is identified and declared, the IPC team implements a set of enhanced infection control measures to attempt to attempt to halt transmission.
The number of cases of infection or colonisation is monitored over a time frame dependent on the pathogen, setting, and scale of the outbreak.
If the measures are successful in controlling the spread, then they are gradually relaxed once rates return to baseline.
If they are not, then further measures are implemented, or the existing measures are intensified, until control is achieved.

In addition to patient factors, the risk of infection and/or colonisation depends on environmental factors (e.g., surface cleaning), proximity to other colonised patients, and use of antimicrobial agents.

@@

The posterior parameter distributions were used to simulate `r length(unique(consts_sim$siminputrow))` outbreak scenarios. Parameter samples were drawn from a uniform distribution:

@@

```{r, include=FALSE}
sens_n_seeds <- length(unique(consts_sim$siminputrow))
```

@@

Using the posterior distributions as inputs to the simulation model, as described in @sec-model-calibration, the model was able to generate scenarios of outbreaks that closely matched the observed data (see @fig-simulated-rates). In these simulated scenarios, the pre-outbreak baseline rate was a median of `r round(pre_outbreak_sims_rates$rate_mean, 2)` cases per 1000 patient days \[interquartile (IQR) range `r round(pre_outbreak_sims_rates$rate_q_low, 2)`--`r round(pre_outbreak_sims_rates$rate_q_high, 2)`\], compared with the real rate of 1.25 cases per 1000 \[IQR 0.94--1.56\]. During the outbreak period, simulations had a mean rate of `r round(outbreak_sims_rates$rate_mean, 2)` \[IQR `r round(outbreak_sims_rates$rate_q_low, 2)`--`r round(outbreak_sims_rates$rate_q_high, 2)`\] cases, compared with the real mean rate of 3.74 \[IQR 3.32--4.41\] (see @fig-simulated-rates).

@@

```{r fig-consts-sims-coloured, include=FALSE}
#| echo: false
#| fig.width: 18
#| fig.height: 8
#| message: false
consts_sim_rates %>% 
  pivot_wider(names_from = date_sim, values_from = rate) %>% 
  #slice_sample(n=10) %>% 
  pivot_longer(cols = !c(`random-seed`, `siminputrow`), names_to = "date_sim", values_to = "rate") %>% 
  mutate(date_sim = ymd(date_sim)) %>% 
  group_by(siminputrow, date_sim) %>%
  summarise(rate = mean(rate)) %>%
  ggplot(aes(x = date_sim,
             y = rate,
             color = factor(`siminputrow`))) +
  geom_line() +
  geom_line(data = long_observed_ss, aes(x = x, y = rates), color = "black") +
  guides(color="none")
```

```{r fig-consts-sims-mean-sd, include=TRUE}
#| echo: false
#| fig-cap: "Simulated rates of _C. difficile_ infection generated by using the posterior parameter distributions. Input parameter samples were generated by sampling the uniform distribution $U_[\\mu - \\sigma, \\mu + \\sigma]$, where $\\mu$ and $\\sigma$ are the mean and standard deviation of the parameter's posterior distribution respectively. The red line shows the true observed infection rates. The black solid line shows the median simulated rates; the dashed lines show the 25^th^ and 75^th^ quantiles."
# plot but with mean and sd instead rather than individual
consts_sim_rates %>% 
  pivot_wider(names_from = date_sim, values_from = rate) %>% 
  pivot_longer(cols = !c(`random-seed`, `siminputrow`), names_to = "date_sim", values_to = "rate") %>% 
  mutate(date_sim = ymd(date_sim)) %>% 
  group_by(date_sim) %>%
  summarise(m = median(rate),
            q_high = quantile(rate, 0.75),
            q_low = quantile(rate, 0.25),
            .groups = "keep") %>%
  ggplot() +
  geom_line(aes(x = date_sim, y = m)) +
  geom_line(aes(x = date_sim, y = q_high), linetype='dashed') +
  geom_line(aes(x = date_sim, y = q_low), linetype='dashed') + 
  geom_line(data = long_observed_ss, aes(x = x, y = rates), color = "red") +
  guides(color="none")
```

>>>>

In practice, the approach to managing outbreaks is often reactive.
Once an outbreak is identified and declared, the IPC team implements a set of enhanced infection control measures to attempt to attempt to halt transmission.
The number of cases of infection or colonisation is monitored over a time frame dependent on the pathogen, setting, and scale of the outbreak.
If the measures are successful in controlling the spread, then they are gradually relaxed once rates return to baseline.
If they are not, then further measures are implemented, or the existing measures are intensified, until control is achieved.
A inherent limitation of this approach is that decisions are made based on observed data, and the impact of interventions cannot be quantified until after a change in the rate of infection has been observed, by which time further preventable nosocomial infections may have occurred.

Computer simulation approaches such as agent-based modelling (ABM) can be used to simulate the spread of infectious diseases by constructing a virtual environment (such as a hospital), within which agents (such as patients and/or staff) interact. Agent-based modelling has been used to model various infectious diseases, such as the transmission of important nosocomial pathogens such as methicillin-resistant _Staphylococcus aureus_ (MRSA) within a healthcare setting [@jaramilloAgentBasedModel2015], and the spread of SARS-CoV-2 within a community [@kerrCovasimAgentbasedModel2021].

@@

The posterior parameter distributions were used to simulate `r length(unique(uniform_sims$siminputrow))` outbreak scenarios. Parameter samples were drawn from a uniform distribution:

@@

```{r, include=FALSE}
sens_n_seeds <- length(unique(uniform_sims$siminputrow))
```

@@

Using the posterior distributions as inputs to the simulation model, as described in @sec-model-calibration, the model was able to generate scenarios of outbreaks that closely matched the observed data (see @fig-posterior-predictive-check). In these simulated scenarios, the pre-outbreak baseline rate was a median of `r round(pre_outbreak_uniform_sims_rates$rate_mean, 2)` cases per 1000 patient days \[interquartile (IQR) range `r round(pre_outbreak_uniform_sims_rates$rate_q_low, 2)`--`r round(pre_outbreak_uniform_sims_rates$rate_q_high, 2)`\], compared with the real rate of 1.25 cases per 1000 \[IQR 0.94--1.56\]. During the outbreak period, simulations had a mean rate of `r round(outbreak_sims_rates$rate_mean, 2)` \[IQR `r round(outbreak_sims_rates$rate_q_low, 2)`--`r round(outbreak_sims_rates$rate_q_high, 2)`\] cases, compared with the real mean rate of 3.74 \[IQR 3.32--4.41\] (see @fig-simulated-rates).

@@

```{r fig-consts-sims-coloured, include=FALSE}
#| echo: false
#| fig.width: 18
#| fig.height: 8
#| message: false
uniform_sims_rates %>% 
  pivot_wider(names_from = date_sim, values_from = rate) %>% 
  #slice_sample(n=10) %>% 
  pivot_longer(cols = !c(`random-seed`, `siminputrow`), names_to = "date_sim", values_to = "rate") %>% 
  mutate(date_sim = ymd(date_sim)) %>% 
  group_by(siminputrow, date_sim) %>%
  summarise(rate = mean(rate)) %>%
  ggplot(aes(x = date_sim,
             y = rate,
             color = factor(`siminputrow`))) +
  geom_line() +
  geom_line(data = long_observed_ss, aes(x = x, y = rates), color = "black") +
  guides(color="none")
```

```{r fig-consts-sims-mean-sd, include=TRUE}
#| echo: false
#| fig-cap: "Simulated rates of _C. difficile_ infection generated by using the posterior parameter distributions. Input parameter samples were generated by sampling the uniform distribution $U_[\\mu - \\sigma, \\mu + \\sigma]$, where $\\mu$ and $\\sigma$ are the mean and standard deviation of the parameter's posterior distribution respectively. The red line shows the true observed infection rates. The black solid line shows the median simulated rates; the dashed lines show the 25^th^ and 75^th^ quantiles."
# plot but with mean and sd instead rather than individual
uniform_sims_rates %>% 
  pivot_wider(names_from = date_sim, values_from = rate) %>% 
  pivot_longer(cols = !c(`random-seed`, `siminputrow`), names_to = "date_sim", values_to = "rate") %>% 
  mutate(date_sim = ymd(date_sim)) %>% 
  group_by(date_sim) %>%
  summarise(m = median(rate),
            q_high = quantile(rate, 0.75),
            q_low = quantile(rate, 0.25),
            .groups = "keep") %>%
  ggplot() +
  geom_line(aes(x = date_sim, y = m)) +
  geom_line(aes(x = date_sim, y = q_high), linetype='dashed') +
  geom_line(aes(x = date_sim, y = q_low), linetype='dashed') + 
  geom_line(data = long_observed_ss, aes(x = x, y = rates), color = "red") +
  guides(color="none")
```
